---
title: 'crasch'
author: "Amy E. Arneson"
date: "1 November 2016"
output: ioslides_presentation
---

```{r setup, include = FALSE}

```

# Introduction

## Objectives

- Learn how to obtain the development version of `crasch`
- View "standard output" provided by the package
- Provide feedback and suggestions for future development

## Basics

- What is `crasch`?

    + `crasch` is an `R` package written to accompany *Constructing Measures* 
      (Wilson, 2005).
    + It is also the (future) reporting engine for the Berkeley Assessment 
      System Software [http://bass.bearapps.org/](http://bass.bearapps.org/).

- What is an `R` package?

    + Freely available code for use in `R` written, maintained, and made 
      available by an `R` user.

- What is `R`?

    + `R` is free statistical software (Mac, PC, or Unix) for computing and 
       graphic production.
    + [https://www.r-project.org](https://www.r-project.org)

## IRT in `R`

Item response theory (IRT) requires specialized statistical analyses, and these 
types of functions are not included in the base `R` libraries.

The following packages are well-documented and widely used for IRT analyses.

- mirt (Chalmers, 2012)
- **TAM** (Kiefer, Robitzsch, \& Wu, 2015)
- eRm (Mair, Hatzinger, \& Maier, 2015)
- ltm (Rizopoulos, 2006)
- **WrightMap** (Torres Irribarra \& Freund, 2014)

## `R` and RStudio

- Download and install `R`:
    + OSX: [http://cran.cnr.berkeley.edu/bin/macosx/](http://cran.cnr.berkeley.edu/bin/macosx/)
    + Win: [http://cran.cnr.berkeley.edu/bin/windows/base/](http://cran.cnr.berkeley.edu/bin/windows/base/)
    + Linux: [http://cran.cnr.berkeley.edu/bin/linux/](http://cran.cnr.berkeley.edu/bin/linux/)

- Download and install RStudio (for a much friendlier `R` experience!)
    + [https://www.rstudio.com/products/rstudio/download/](https://www.rstudio.com/products/rstudio/download/)
    + Choose your OS under "Installers" and proceed as you would with any 
      software installation.
    
- If you already have `R` on your machine, you may need to update to the latest 
  version.  Use the links above to do so.

## Demo materials

- These slides, the associated RMarkdown file, and a script with all commands
  are available at [https://github.com/amyarneson/crasch-materials/BEARseminar](https://github.com/amyarneson/crasch-materials/BEARseminar)
  
## Installing `crasch`

```{r, eval = FALSE}
# This first package allows you to install from github
install.packages("devtools")

# Install crasch
devtools::install_github("amyarneson/crasch")
# This may take a few minutes; multiple packages will be installed
# You may see some warning messages, and that's okay!

# Use the same command to update the package as changes are made.
```

# How it works

## Input files

## Installing/Updating the package

## Current capabilities

## Future plans

# Standard analysis: Demo

## Estimation

## "Standard" output

# Validity

## Internal structure

### Instrument level

Wright Map arranged by theorized construct levels

### Item level

Mean locations

### Item by group level

DIF - not available yet :/

## External relationships
