---
title: 'crasch'
author: "Amy E. Arneson"
date: "1 November 2016"
output: ioslides_presentation
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)

library(crasch)
```

# Introduction

## Objectives

- Learn how to obtain the development version of `crasch`
- View "standard output" provided by the package
- Provide feedback and suggestions for future development

## Basics

- What is `crasch`?

    + `crasch` is an `R` package written to accompany *Constructing Measures* 
      (Wilson, 2005).
    + It is also the (future) reporting engine for the Berkeley Assessment 
      System Software [http://bass.bearapps.org/](http://bass.bearapps.org/).

- What is an `R` package?

    + Freely available code for use in `R` written, maintained, and made 
      available by an `R` user.

- What is `R`?

    + `R` is free statistical software (Mac, PC, or Unix) for computing and 
       graphic production.
    + [https://www.r-project.org](https://www.r-project.org)

## IRT in `R`

Item response theory (IRT) requires specialized statistical analyses, and these 
types of functions are not included in the base `R` libraries.

The following packages are well-documented and widely used for IRT analyses.

- mirt (Chalmers, 2012)
- **TAM** (Kiefer, Robitzsch, \& Wu, 2015)
- eRm (Mair, Hatzinger, \& Maier, 2015)
- ltm (Rizopoulos, 2006)
- **WrightMap** (Torres Irribarra \& Freund, 2014)

## `R` and RStudio

- Download and install `R`:
    + OSX: [http://cran.cnr.berkeley.edu/bin/macosx/](http://cran.cnr.berkeley.edu/bin/macosx/)
    + Win: [http://cran.cnr.berkeley.edu/bin/windows/base/](http://cran.cnr.berkeley.edu/bin/windows/base/)
    + Linux: [http://cran.cnr.berkeley.edu/bin/linux/](http://cran.cnr.berkeley.edu/bin/linux/)

- Download and install RStudio (for a much friendlier `R` experience!)
    + [https://www.rstudio.com/products/rstudio/download/](https://www.rstudio.com/products/rstudio/download/)
    + Choose your OS under "Installers" and proceed as you would with any 
      software installation.
    
- If you already have `R` on your machine, you may need to update to the latest 
  version.  Use the links above to do so.

## Demo materials

- These slides, the associated RMarkdown file, and a script with all commands
  are available at [https://github.com/amyarneson/BEARseminar](https://github.com/amyarneson/BEARseminar)
    
    + Github will render the `.Rmd` file and you can view these slides on github
      (just click on the `.Rmd` file!).
    + You can also view the R script (the `.R` file) on github.
    + To download materials: click the green "Clone or download" button and 
      click "Download ZIP."
  
## Installing `crasch`

```{r, eval = FALSE}
# This first package allows you to install from github
install.packages("devtools")

# Install crasch
devtools::install_github("amyarneson/crasch")
# This may take a few minutes; multiple packages will be installed
# You may see some warning messages, and that's okay!

# Use the same command to update the package as changes are made.
```

# How it works

## Input files

There are 3 required input files to run your analysis and an optional 4th:
  
- construct information (`consInfo`)
- item information (`itemInfo`)
- scored responses (`scores`)
- external variables (optional, `varsInfo`)

Each of these files must structured exactly as shown in this tutorial with the
  correct column headings.  Sample data files are available at
  [https://github.com/amyarneson/crasch-materials](https://github.com/amyarneson/crasch-materials)
  
*I will not be covering data files in detail today.  The above link has a 
  tutorial devoted only to making the data files.*
  
## Input files, cont.

The package comes with sample data sets.

The ADP data corresponds with a construct developed for a study of teacher 
  adoption of a cross-curricular writing "program" in a Missouri public high 
  school in 2013.
  
    + *A PDF with some more information about the construct and instrument is 
      provided on github.*
      
    + It was Amy's 274A class project.

```{r inputFiles, eval = FALSE}
library(crasch)

# You can view the input objects
View(ADPcons) # Construct information input
View(ADPitem) # Item information input
View(ADPwide) # Scores in wide format
```

# Standard analysis: Demo

## Estimation

The main function for the package is `craschR()`.

- This function calls `TAM` to perform estimation and then organizes the output
  in a friendly way.
  
- Output can be saved to your working directory in CSV format using 
  `writeout = TRUE`.
  
- Save the results as an R object so that they can be used in the 
  post-estimation functions.
  
```{r analysis, eval = FALSE, message = FALSE}
?craschR  # check out the argument options and description

ADP = craschR(scores = ADPwide, itemInfo = ADPitem, 
              consInfo = ADPcons, varsInfo = ADPvars, 
              writeout = FALSE)
```

```{r analysis0, include= FALSE}
ADP = craschR(scores = ADPwide, itemInfo = ADPitem, 
              consInfo = ADPcons, varsInfo = ADPvars, 
              writeout = FALSE)
```

## Estimation

The default person estimation method is EAP.

You can also obtain MLEs or WLEs.

```{r personOptions, eval = FALSE}
ADP_MLE = craschR(scores = ADPwide, itemInfo = ADPitem, 
                  consInfo = ADPcons, varsInfo = ADPvars, 
                  persMethod = "MLE", writeout = FALSE)
ADP_WLE = craschR(scores = ADPwide, itemInfo = ADPitem, 
                  consInfo = ADPcons, varsInfo = ADPvars, 
                  persMethod = "WLE", writeout = FALSE)

# View the different estimates in a table
cbind(ADP$persPars, ADP_MLE$persPars, ADP_WLE$persPars)
```

## Estimation

To access the output in your R session, use `$` with the name of the object you
  want to view.

```{r names}
names(ADP)
```

See the `craschR()` help documentation for information on the format of each
  element of the list.

```{r viewingOutput, eval = FALSE}
View(ADP$itemPars)
View(ADP$persPars)
ADP$estSummary
```

# Post-Estimation functions

## `item.scores()`
### Produce tables and graphs for the score distribution by item

```{r item.scores, message = FALSE, results = "hide"}
?item.scores
item.scores(ADP)
```

## `item.scores()` cont.

```{r item.scores2, message = FALSE, results = "hide"}
# To graph proportions instead of frequencies, use freqs = FALSE
item.scores(ADP, freqs = FALSE)
```

## `item.scores()` cont.

Check out the Color Brewer website.  It is really helpful in determining color
palettes in general for presentations.  They even have color-blind-friendly
palettes if you want to make a fully-accessible presentation!

[colorbrewer2.org](colorbrewer2.org)

The `RColorBrewer` package pulls this information from the Color Brewer website.
Each color scheme has a name.  You can then put that scheme name into the
`palette` argument in a `crasch` graphic function.

## `item.scores()` cont.

```{r colorBrew, message = FALSE, results = "hide"}
display.brewer.all() # creates a graphic with the palettes displayed + names
```

## `item.scores()` cont.

```{r colorBrewOpts, message = FALSE, results = "hide"}
item.scores(ADP, palette = "Reds")
```

## `item.scores()` cont.
### Graphic file writeout options

- PDF is the default file type, but can be difficult to work with in MS Office.

- SVGs are also supported (note that extra software may be required).

- PNGs, JPEGs, BMPs, and TIFFs are lower-quality graphics, but are easily 
  inserted in MS doc & ppt files.


```{r item.scoresWrite1, eval = FALSE}
item.scores(ADP, writeout = TRUE, imageType = "png")
item.scores(ADP, writeout = TRUE, imageType = "jpeg")
item.scores(ADP, writeout = TRUE, imageType = "bmp")
item.scores(ADP, writeout = TRUE, imageType = "tiff")
```

## `pers.hist()`
### Estimated person locations with a normal curve overlaid

```{r pers.hist, results = "hide", message = FALSE}
?pers.hist
pers.hist(ADP)
```

## `pers.hist()` cont.

You can change the colors and the write options as shown before.

```{r pers.hist2, eval = FALSE}
pers.hist(ADP, palette = "Spectral")
# You can also specify any 2 R colors in a vector
  pers.hist(ADP, palette = c("red", "darkred"))
```

For a list of R color names: [http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf)

## `KIDMAP()`
### Produce a KIDMAP graphic for a given respondent

```{r KIDMAP, message = FALSE}
?KIDMAP
# For the KIDMAP function you must specify a respondent ID!
# To look at the list of your IDs:
row.names(ADPwide)
```

## `KIDMAP()` cont.

```{r KIDMAP2}
KIDMAP(ADP, 120)
```

## `KIDMAP()` cont.

You can change the colors and the write options as above.

```{r KIDMAP3, eval = FALSE}
KIDMAP(ADP, 120, palette = "Spectral")
# You can also specify any 3 R colors in a vector
  KIDMAP(ADP, 147, palette = c("darkred", "red", "black"))
```

## `item.analysis()`
### Produce by item and by step tables

```{r item.analysis, message = FALSE, eval = FALSE}
?item.analysis

item.analysis(ADP)
```

Truncated output (item table):

```{r item.analysis2, echo = FALSE, message = FALSE}
item.analysis(ADP)[[1]][1:3,]
```

## `item.analysis()` cont.

Truncated output (step table): 

```{r item.analysis3, echo = FALSE, message = FALSE}
item.analysis(ADP)[[2]][1:3,]
```

## `infit.MNSQ()`
### Produce a graphic of item fit statistics

```{r infitMNSQ, message = FALSE}
?infit.MNSQ
infit.MNSQ(ADP)
```

## `infit.MNSQ()` cont.

To plot the step fit instead of the overall item fit:
```{r infitMNSQ2, message = FALSE}
infit.MNSQ(ADP, params = "steps")
```

## `infit.MNSQ()` cont.

To restrict which items are plotted on each graph:
```{r infitMNSQ3, message = FALSE}
infit.MNSQ(ADP, itemOrder = c(1, 2, 3, 4, 5, 6, 7), params = "steps")
infit.MNSQ(ADP, itemOrder = 8:13, params = "steps")
```

## `CPC.graph()`
### Produce graphics of cumulative probability curves

```{r CPC.graph, eval = FALSE}
?CPC.graph
CPC.graph(ADP) # Note that a separate graph is produced for each item
```

## `CPC.graph()` cont.

To selectively produce graphs use itemOrder:
```{r CPC.graph2, message = FALSE}
CPC.graph(ADP, itemOrder = c(1, 13))
```

## `CPC.graph()` cont.

To plot your observed proportions, use observed and minCell:
```{r CPC.graph3, message = FALSE, warning = FALSE}
CPC.graph(ADP, itemOrder = 13, observed = TRUE)
```

## `CPC.graph()` cont.

```{r CPC.graph4, message = FALSE}
CPC.graph(ADP, itemOrder = 13, observed = TRUE, minCell = 3)
```

## `CPC.graph()` cont.

You can also plot "vertically" as in Wilson (2005)!

```{r CPC.graph5, message = FALSE}
CPC.graph(ADP, itemOrder = 13, logitAxis = "y")
```

## `ICC.graph()`
### Produce graphics of the item (or category) characteristic curves

```{r ICC.graph, eval = FALSE}
?ICC.graph

ICC.graph(ADP) # produces a graph for each item
```

## `ICC.graph()` cont.

itemOrder works as in the CPC.graph() function
```{r ICC.graph2, message = FALSE}
ICC.graph(ADP, itemOrder = c(5:6))
```

## `info.graph()`
### Produce graphics of test and item information

```{r info.graph, message = FALSE}
?info.graph
info.graph(ADP, type = "SEM")  # SEM is the default
```

## `info.graph()` cont.

You can plot all observations (even incomplete ones) using 
  `completeOnly = FALSE`
  
```{r info.graph2, message = FALSE}
info.graph(ADP, type = "SEM", completeOnly = FALSE)

```

## `info.graph()` cont.

For the test information curve:

```{r info.graph3, message = FALSE}
info.graph(ADP, type = "TIC")  # produces one graph
```

## `info.graph()` cont.

For the item information curves: 

```{r info.graph4, message = FALSE}
info.graph(ADP, type = "IIC")  # produces 1 graph per item!
```

## `wm()`
### A wrapper for the `wrightMap()` function

```{r wm, message = FALSE}
?wm
wm(ADP) # organized in item order
```

## `wm()` cont.

```{r wm2, message = FALSE}
wm(ADP, byCat = TRUE) # organized in construct category order
```

## `wm()` cont.

If you want to customize the output further, see the `wrightMap()` function 
  documentation.  The function provided in this package is merely a wrapper.
  
```{r wrightMap, eval = FALSE}
?wrightMap
```

## `Split.halves()`
### Calculate the split halves reliability coefficient

There is some data preprocessing that needs to happen before runnning 
  `Split.halves()`.
  
1. You must split your item information and score objects in half.

2. You must run `craschR()` on each half.

```{r Split.halves, message = FALSE}
item1 = ADPitem[c(2, 5, 7, 8, 10, 11), ]      # selects rows
scores1 = ADPwide[ , c(2, 5, 7, 8, 10, 11)] # selects columns

item2 = ADPitem[c(1, 3, 4, 6, 9, 12, 13), ]
scores2 = ADPwide[ , c(1, 3, 4, 6, 9, 12, 13)]
```

## `Split.halves()` cont.

```{r Split.halves2, message = FALSE}
results1 = craschR(scores1, item1, ADPcons, writeout = FALSE)
results2 = craschR(scores2, item2, ADPcons, writeout = FALSE)

?Split.halves

Split.halves(results1, results2)
```

## `mn.traj()`
### Graph showing mean location trajectories by item

```{r mn.traj, message = FALSE}
?mn.traj
mn.traj(ADP)
```

## `Sp.rho()`
### Calculate Spearman's rho

We need fully dichotomous data for this function.  A sample data set is provided
  in the package.

```{r Sp.rho, message = FALSE}
?Sp.rho

DIresults = craschR(scores = DIwide, itemInfo = DIitem, 
                    consInfo = DIcons, writeout = FALSE)

Sp.rho(DIresults, expOrder = c(14, 13, 12, 11, 10, 9, 8, 7, 
                               6, 5, 4, 3, 2, 1))
```

# Future development

## Future plans

- Write estimation code instead of calling other packages?

- Implement in BASS for reporting

- `mirt` support

- More flexibility/options in available functions:
    + Rating scale model
    + Graphic options (bin size, etc.)

- New functions:
    + DIF
    + Import results from ConQuest, Stata, or other R packages
    
# Thanks!

aea@berkeley.edu

Thanks to package contributors: David Torres Irribarra, Rebecca Freund, Yukie
  Toyama, and Mark Wilson

bearcenter.berkeley.edu